### 2024-07-17 Article/Article-english.tex
translation

\medskip\noindent\textbf{Теорема} (Шмидт) \cite{Schm94}. \emph{Let $K$ be a~field. Every algebraic subgroup $H$
    of the~group $(K^{\times})^n$ is~defined by~a~system of~some number $N$ of binomial equations, namely,
    there are $N$ indices $\alpha_i, \beta_i \in \mathbb{Z}^n$ such that
    $$
        H = \{ z \in (K^{\times})^n\ |\ \forall 1 \leq i \leq N{:}\ z^{\alpha_i} = z^{\beta_i} \}.
    $$
}

@@

\noindent\textbf{Теорема} (Артин) \cite{Art48}.\emph{
    Any $n$ pairwise distinct characters are linearly independent.
}\medskip

@@

\medskip\noindent\textbf{Theorem~1.}\emph{
    Let $\alpha \in \mathbb{Z}^{n \times k}$. The~following statements are equivalent:
    \begin{enumerate}
        \item The~rows of the~matrix $\alpha$ span the~whole lattice $\mathbb{Z}^k$.
        \item The~minors of the~maximum order of the~matrix $\alpha$ are coprime.
        \item $\mathrm{SNF}(\alpha) = \mathrm{diag}^{n \times k}_k(1, \ldots, 1)$.
    \end{enumerate}
}

@@

\medskip\noindent\textbf{Теорема 3.}\emph{
    The~number of connected components of $\ker(\phi_\beta) \subseteq ({\mathbb{C}^\times})^n$ is equal to~$\Pi(\beta)$.
}\medskip

@@

stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub
stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub
stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub
stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub
stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub
stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub
stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub
stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub
stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub
stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub stub

>>>>

\medskip\noindent\textbf{Theorem} (Schmidt) \cite{Schm94}. \emph{Let $K$ be a~field. Every algebraic subgroup $H$
    of the~group $(K^{\times})^n$ is~defined by~a~system of~some number $N$ of binomial equations, namely,
    there are $N$ indices $\alpha_i, \beta_i \in \mathbb{Z}^n$ such that
    $$
        H = \{ z \in (K^{\times})^n\ |\ \forall 1 \leq i \leq N{:}\ z^{\alpha_i} = z^{\beta_i} \}.
    $$
}

@@

\noindent\textbf{Theorem} (Artin) \cite{Art48}.\emph{
    Any $n$ pairwise distinct characters are linearly independent.
}\medskip

@@

\medskip\noindent\textbf{Theorem~1.}\emph{
    Let $\alpha \in \mathbb{Z}^{n \times k}$. The~following statements are equivalent:
    \begin{enumerate}
        \item The~rows of the~matrix $\alpha$ span the~whole lattice $\mathbb{Z}^k$.
        \item The~maximal minors of the~matrix $\alpha$ are coprime.
        \item $\mathrm{SNF}(\alpha) = \mathrm{diag}^{n \times k}_k(1, \ldots, 1)$.
    \end{enumerate}
}

@@

\medskip\noindent\textbf{Theorem~3.}\emph{
    The~number of connected components of $\ker(\phi_\beta) \subseteq ({\mathbb{C}^\times})^n$ is equal to~$\Pi(\beta)$.
}\medskip

@@

\noindent{\it Proof.}
    The~left-to-right implication is obvious. Conversely, let $\mu_1, \ldots, \mu_n$ be linearly independent over $\mathbb{Q}$.
    Consider their linear combination: $r_1 \mu_1 + \ldots + r_n \mu_n = 0$.

    It is known that $\mathbb{R}$ is an~(infinite-dimensional) vector space over $\mathbb{Q}$, and any vector space has
    a~(Hamel) basis under assumption of the~axiom of~choice~\cite{Brbk70}. Using this, let us fix a~Hamel basis $B$ for $\mathbb{R}$
    over $\mathbb{Q}$. Let us decompose $r_i$ by this basis:
    $$
        r_i = q_i^1 b_1 + \ldots + q_i^s b_s
    $$
    where $q_i^j \in \mathbb{Q}$ and $b_j \in B$. There are finitely many vectors $r_i$,
    so the~set of basis vectors $b_j$ for them can be chosen to be the~same.
    Let us substitute the~decomposition into the~linear combination:
    \begin{align*}
        r_1 \mu_1 + \ldots + r_n \mu_n & = (q_1^1 b_1 + \ldots + q_1^s b_s) \mu_1 + \ldots + (q_n^1 b_1 + \ldots + q_n^s b_s) \mu_n \\
                                       & = (q_1^1 \mu_1 + \ldots + q_n^1 \mu_n) b_1 + \ldots + (q_1^s \mu_1 + \ldots + q_n^s \mu_n) b_s \\
                                       & = 0.
    \end{align*}

    In each coordinate we obtain a~rational linear combination of the~numbers $b_j$ equal to zero.
    By~virtue of linear independence, all these coordinates are equal to zero, so we have:
    $q_1^j \mu_1 + \ldots + q_n^j \mu_n = 0$.

    However, all $q_i^j$ are rational, and $\mu_i$ are linearly independent over $\mathbb{Q}$, so $q_i^j = 0$.
    Thus, $r_i = q_i^1 b_1 + \ldots + q_i^s b_s = 0 \cdot b_1 + \ldots + 0 \cdot b_s = 0$.
\hfill$\Box$\medskip

\medskip\noindent\textbf{Lemma.}\emph{
    $\mu_1, \ldots, \mu_n \in \mathbb{Z}^k$ are linearly independent over $\mathbb{Q}$ if and only if they are linearly independent over $\mathbb{Z}$.
}\medskip

\noindent{\it Proof.}
    The~left-to-right implication is again obvious. Conversely, let $q_1 \mu_1 + \ldots + q_n \mu_n = 0$ where $q_i \in \mathbb{Q}$.
    We choose the~common denominator $q \in \mathbb{N} \setminus \{ 0 \}$ for the~fractions $q_i$.
    We denote the~numerators by $p_i \in \mathbb{Z}$, that is, $q_i = p_i / q$.

    Then $(p_1 \mu_1 + \ldots + p_n \mu_n) / q = 0$, so $p_1 \mu_1 + \ldots + p_n \mu_n = 0$.
    All $p_i$ are integers, so using linear independence over $\mathbb{Z}$ we obtain that $p_1 = \ldots = p_n = 0$.
    Thus, $q_i = p_i / q = 0 / q = 0$.
\hfill$\Box$\medskip

\medskip\noindent\textbf{Consequence.}\emph{
    $\mu_1, \ldots, \mu_n \in \mathbb{Z}^k$ are linearly independent over $\mathbb{R}$ if and only if they are linearly independent over $\mathbb{Z}$.
}\medskip

Let us study in~more detail the~corollaries of linear independence over $\mathbb{R} / 2 \pi \mathbb{Z}$.

\medskip\noindent\textbf{Proposition.}\emph{
    $\mu_1, \ldots, \mu_n$ are linearly independent over $\mathbb{R} / 2 \pi \mathbb{Z}$
    if and only if they are linearly independent over $\mathbb{R} / \mathbb{Z}$.
}\medskip

\noindent{\it Proof.}
    This follows from that the~groups $\mathbb{R} / 2 \pi \mathbb{Z}$ and $\mathbb{R} / \mathbb{Z}$ are isomorphic.
\hfill$\Box$\medskip

For a~number $d \in \mathbb{Z}$ and a~vector $x \in \mathbb{Z}^k$ by $d \mid x$ we mean that $d \mid x_i$ for all $1 \leq i \leq k$
or, equivalently, $d \mid \gcd(x_1, \ldots, x_n)$.

\medskip\noindent\textbf{Lemma.}\emph{
    If $\mu_1, \ldots, \mu_n$ are linearly independent over $\mathbb{R} / \mathbb{Z}$, then:
    $$
        \forall d \in \mathbb{Z}{:}\ \forall x \in \mathbb{Z}^n{:}\ d \mid x_1 \mu_1 + \ldots + x_n \mu_n \Rightarrow d \mid x.
    $$
}

\noindent{\it Proof.}
    Indeed, consider $q_i = x_i / d \in \mathbb{R} / \mathbb{Z}$. The~divisibility of $x_1 \mu_1 + \ldots + x_n \mu_n$ by~$d$ means
    that in~$\mathbb{R} / \mathbb{Z}$ it holds that $(x_1 \mu_1 + \ldots + x_n \mu_n) / d = 0$, i.e. $q_1 \mu_1 + \ldots + q_n \mu_n = 0$.
    However, by virtue of linear independence, it is true that $q_1 = \ldots = q_n = 0$, but this means $d \mid x_i$ for all $1 \leq i \leq n$.
\hfill$\Box$\medskip

A~simple non-injectivity criterion follows immediately.

\medskip\noindent\textbf{Lemma.}\emph{
    If $\phi_\alpha$ is injective, then $\forall j{:}\ \gcd(\alpha^j) = 1$.
}

\medskip\noindent\textbf{Consequence.}\emph{
    If $\exists j{:}\ \gcd(\alpha^j) \neq 1$, then $\phi_\alpha$ is not an~injective mapping.
}\medskip

\noindent{\it Proof.}
    Choosing $x_j = 1$ and $x_1 = \ldots = x_{j - 1} = x_{j + 1} = \ldots = x_k = 0$, we obtain that:
    \begin{align*}
                         & \forall d \in \mathbb{Z}{:}\ \forall x \in \mathbb{Z}^k{:}\ d \mid x_1 \alpha^1 + \ldots + x_k \alpha^k \Rightarrow d \mid x \\
            \Rightarrow\ & \forall j{:}\ \forall d \in \mathbb{Z}{:}\ d \mid \alpha^j \Rightarrow d \mid 1
        \Leftrightarrow \forall j{:}\ \gcd(\alpha^j) = 1.
        \tag*{$\Box$}
    \end{align*}
{}

We denote the~set of prime numbers by $\mathbb{P} \subseteq \mathbb{Z}$.

\medskip\noindent\textbf{Lemma.}\emph{
    For arbitrary integer vectors $\mu_1, \ldots, \mu_n$ it is true that:
    \begin{align*}
                         & \forall d \in \mathbb{Z}{:}\ \forall x \in \mathbb{Z}^n{:}\ d \mid x_1 \mu_1 + \ldots + x_n \mu_n \Rightarrow d \mid x \\
        \Leftrightarrow\ & \forall p \in \mathbb{P}{:}\ \forall \beta \in \mathbb{N}{:}\ \forall x \in \mathbb{Z}^n{:}\ p^\beta \mid x_1 \mu_1 + \ldots + x_n \mu_n \Rightarrow p^\beta \mid x \\
        \Leftrightarrow\ & \forall p \in \mathbb{P}{:}\ \forall x \in \mathbb{Z}^n{:}\ p \mid x_1 \mu_1 + \ldots + x_n \mu_n \Rightarrow p \mid x
    \end{align*}
}

\noindent{\it Proof.}
    The left-to-right implications are obvious. Let us prove the~right-to-left ones.
    Fix an~arbitrary $d \in \mathbb{Z}$ and some vector $x \in \mathbb{Z}^k$.
    Factor $d$ into prime numbers: $d = p_1^{\beta_1} \ldots p_m^{\beta_m}$.

    Since $x_1 \mu_1 + \ldots + x_n \mu_n$ is divisible by~$d$, it is also divisible by every $p_i^{\beta_i}$.
    Using premise, we obtain that $p_i^{\beta_i} \mid x$ for all $1 \leq i \leq m$; but then
    $d = p_1^{\beta_1} \ldots p_m^{\beta_m} \mid x$ as required.

    Next, fix a~power $\beta \in \mathbb{N}$ and a~prime number $p$.
    Since $x_1 \mu_1 + \ldots + x_n \mu_n$ is divisible $p^\beta$,
    it is also divisible by~$p$, so it follow from the~premise that $p \mid x$.

    But this means that $x_i / p$ are integers, so $(x_1 / p) \mu_1 + \ldots + (x_n / p) \mu_n$ is divisible by~$p^{\beta - 1}$.
    Applying the~premise again, we obtain that $p \mid x / p$. Repeating this procedure $\beta$ times, we finally conclude
    that $p \mid x / p^{\beta - 1}$; but this is equivalent to~$p^\beta \mid x$.
\hfill$\Box$\medskip

The last condition can be rewritten as linear independence of vectors $\mu_i$ over the~fields $\mathbb{Z} / p\mathbb{Z}$
for all primes $p$, which is equivalent to having the~full rank for the~matrix $\mu$.

\medskip\noindent\textbf{Consequence.}\emph{
    For arbitrary integer vectors $\mu_1, \ldots, \mu_n$ it is true that:
    \begin{align*}
                         & \forall d \in \mathbb{Z}{:}\ \forall x \in \mathbb{Z}^n{:}\ d \mid x_1 \mu_1 + \ldots + x_n \mu_n \Rightarrow d \mid x \\
        \Leftrightarrow\ & \forall p \in \mathbb{P}{:}\ \forall x \in (\mathbb{Z} / p\mathbb{Z})^n{:}\ x_1 \mu_1 + \ldots + x_n \mu_n = 0 \Rightarrow x = 0 \\
        \Leftrightarrow\ & \forall p \in \mathbb{P}{:}\ \mathrm{rank}_{\mathbb{Z} / p \mathbb{Z}}(\mu) = \max(n, k).
    \end{align*}
}

Finally, we can re-prove the~already mentioned sufficient condition.

\medskip\noindent\textbf{Consequence.}\emph{
    If $\phi_\alpha$ is injective, then $\alpha_i$ span the~whole lattice.
}\medskip

\noindent{\it Proof.}
    Assume that $\alpha_i$ do not generate the~lattice. Then the~maximal minors of the~matrix $\alpha$ have a~common divisor $d > 1$.

    Take some prime divisor $p$ of the~number $d$. Since the~maximal minors are divisible by~$d$, they are also divisible by $p$;
    therefore in the~field $\mathbb{Z} / p \mathbb{Z}$ all maximal minors of $\alpha$ are equal to zero.
    So we have that $\mathrm{rank}_{\mathbb{Z} / p \mathbb{Z}}(\alpha) < \max(n, k)$~\cite{Brbk70},
    but this contradicts the~corollary.
\hfill$\Box$
